---
title: "confidence-intervall"
author: "Nicolas Stern"
date: "27 décembre 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Calculating a confidence intervall

cf https://www.cyclismo.org/tutorial/R/confidence.html#id2

## Calculating a confidence intervall from a normal distribution
Here we will look at a fictitious example. We will make some assumptions for what we might find in an experiment and find the resulting confidence interval using a normal distribution.

Let's assume :

- sample mean is 5

- the standard deviation is 2

- the sample size is 20. 

In the example below w"e will use a 95% confidence level and wish to find the confidence interval. 

The commands to find the confidence interval in R are the following:
```{r}
a <- 5
s <- 2
n <- 20
error <- qnorm(0.975)*s/sqrt(n)
left <- a-error
right <- a+error
left
right
```
## Calculating a confidence intervall from a t distribution

```{r}
a <- 5
s <- 2
n <- 20
error <- qt(0.975,df=n-1)*s/sqrt(n)
left <- a-error
right <- a+error
left
right
```

## the easy way: using t.test
```{r}
x = c(9.0,9.5,9.6,10.2,11.6)
t.test(x)
```

```{r}
rm(list = ls())
```

## the example in the exam 2016: confidence interval of a sample
- Die Auswertung einer Stichprobe der Grösse n = 100 ergab ¯ x = 34.25 und s = 7.5. Bestimmen Sie ein 95%-Konﬁdenzintervall für den Populationsmittelwe
```{r}
n <- 100
s <- 7.5
xbar <- 34.25
se <- s/sqrt(n)
alpha <- 0.05

err <- qt( (1-alpha/2), df = n-1) * se
err

int <- c(xbar - err, xbar + err)
int
```


