---
title: "Example Hypothesis Testing"
author: "Nicolas Stern"
date: "30 décembre 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Hypothesis testing

- How to test a two-sided hypothesis based on sample data.
- Source : https://www.youtube.com/watch?v=vw2IPZ2aD-c

- Given the following values, let's calculate the test value using t-statistics to assess our hypothesis:
    - H$_{0}$ :  $\mu$ = 100

    - H$_{a}$ :  $\mu$ != 100

```{r}
xbar <- 104 # sample mean
mu <- 100 # hypothetical population mean
n <- 35 # sample size
s <- 12 # sample standard deviation

# Calculate the test value
t <- (xbar - mu) / (s / sqrt(n))
t
```
We use a t-distribution, NOT the normal distribution, since we are working with sample data. 

T-distribution is different from the normal distribution (although only very slightly). Let's compare both curves. For the t-distribution, we use a value of 34 degrees of freedom (n-1):

```{r}
ndata <- seq(-3, 3, by=0.1)
npoints <- dnorm(ndata)
tpoints <- dt(ndata, df=34)
plot(npoints, type="l", col="blue")
lines(tpoints, col="red")
```

Let's further examine this difference : let's calculate the probabilities using the t-distribution for the given t-value: 

probability using the t-distribution: 

```{r}
pt(t, df=n-1, lower.tail = F) * 2
```
We multiply by 2 to cover the total area under the curve (on the right of the t-value, AND on the left of -t-value).

Compare it with the normal distribution:

```{r}
pnorm(t, lower.tail = F) * 2
```
Now, using the t-distribution, let's examinate our hypothesis. We will reject  H$_{0}$ if the t-value we calculated  (`r t`) is greater than a critical value (on the right side) or smaller (on the left side).

Let's calculate the critical value for a 5 % significance level:
```{r}
alpha <- 0.05
talpha <- qt(1-alpha/2, df = n-1, lower.tail = T)
talpha
```
Since the t-value is smaller than  the critical value, we have no evidence against H$_{0}$.

We could also compare our p-value with the significance level : the greater the p-value, the stronger the evidence against  H$_{0}$. The significance level `r alpha` is a cutoff for significance.

```{r}
p.value <- pt(t, df=24, lower.tail = F) * 2
p.value
p.value > alpha
```
Since the p-value is greater than alpha, the evidence against H$_{0}$ is not significant.